### 2022-11-14 Manuscript/Manuscript.tex
Fixed long lines

\documentclass[12pt]{article}

%% preamble: Keep it clean; only include those you need
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage[table]{xcolor}
\usepackage{booktabs}

% for space filling
% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

%% meta data
\setlength\parindent{24pt}
\linespread{2}
\title{WNBA vs NBA Earnings In Comparable Players}
\author{Colin Sullivan\\
  Department of Statistics\\
  University of Connecticut
}

\begin{document}

   \begin{center}
       \vspace*{1cm}
       \Huge
       \textbf{How to Earn a Salary}

       \Large
       \vspace{0.5cm}
        A Closer Look at WNBA and NBA Player Salaries, the Discrepancies, and What Statistics Might Influence Them

       \vspace{1.5cm}

       \textbf{Colin Sullivan}

       \vfill

       \vfill

       \vspace{0.8cm}

        \Large    
       Department of Statistics\\
       University of Connecticut\\
       Storrs, CT\\

   \end{center}
\newpage

\paragraph{Abstract}
In a world where gender pay gaps have been highlighted in nearly every industry, professional sports are no exception. 
The Women’s National Basketball Association is widely known and accepted to pay very significantly less annually to its 
players than the male league counterpart, the National Basketball Association. There has been a lot of research into this 
pay gap, and this paper will look to add to the conversation by examining the structure of contracts in relation to performance. This will be done using statistics from the 2018 seasons of each respective league, using datasets cleaned up and analyzed using various python scripts to eventually draw conclusions on the relation between performance and pay for male and female athletes.

\paragraph{Introduction}
Since the formulation of the Women’s National Basketball Association (WNBA) in 1996, there have been never-ending debates 
among fans comparing the female counterpart to the much older, more popular National Basketball Association (NBA). 
Women’s basketball has a rich history marred by frequent sexism, fewer opportunities, and significantly less financial support.
\par
Women’s basketball began being played in 1892 at Smith College, just one year after men’s basketball was invented 
\cite{Shattering_The_Glass}. Due to the gender roles and cultural norms, however, women at the time were believed 
to be very fragile and unable to play physical sports like basketball at anywhere near the same level as men and 
thus played under modified rules specially created for them \cite{WNBA_Hist}. Those familiar with basketball will 
know that it is usually played competitively with 5 players on each team, and each player runs the full court to 
play offense on one end and defend the other team on the other end. In 1892, however, women initially played “9 on 9”, 
with the court divided into three areas: offense, defense, and midcourt. Each woman was assigned one of these courts 
in which they played, such that 3 women on each team were in each section. The players could not leave their section 
of the court to go into the other, essentially creating three separate “3 on 3” games. The other major difference 
from the men’s game was that women could only dribble the ball a limited number of times and could not hold it for 
more than three seconds, rule changed attributed to the fear that the women players might suffer “nervous fatigue” 
if the game were too strenuous \cite{WNBA_Hist}.
\par
Everyone now knows that such attitudes towards women’s sports are ridiculous and sexist. Certainly there are some physical 
characteristic differences between men and women in sports, but for every male athlete there is a just as accomplished 
female one. The gradual changing of perspectives on women’s athletics is seen in the evolution of women’s basketball; 
while men’s basketball has largely stayed the same over the course of its almost 130 year history (key changes including 
the legalization of the slam dunk and the implementation of the three point line), women’s basketball has shifted slowly 
to eventually become almost identical. The sport was played inter-collegiately for decades beginning in the late 19th 
century, but did not enter the professional realm until the 1970s. The sport became officially recognized at the collegiate 
level by the NCAA in 1982 (a shockingly late introduction), and the Women’s Basketball League (WBL), founded in 1978 and 
only lasting three seasons, was the first of several failed women’s professional leagues in the late 20th century. In the 
meantime, the NBA experienced a huge boom in popularity largely due to the emergence of superstars Larry Bird, 
Magic Johnson, and Michael Jordan. The massive increase in NBA reach and revenue directly led the league creating, 
funding, and supporting the WNBA in 1996.
\par
There are minimal differences between today's women’s basketball and the basketball that men have been playing since 
inception. The WNBA and NCAAW ball is slightly smaller than that used by male players, the three point line is slightly 
shorter, and the game clock features four 10 minute quarters as opposed to the NBA’s 12 minute quarters, but for all 
intents and purposes the games are the same. The rules are otherwise identical, the athletes are similarly talented, 
the level of play is incredibly high, and the games are very enjoyable to watch. There remains only one massive 
difference between the two leagues: player salaries.
\par
The difference in player salaries is a huge point of contention in women’s basketball discourse and is often the subject 
of intense ridicule, and it is easy to see why: top NBA players make upwards of 40 million USD per season, not including 
endorsement deals and other income, while the WNBA most valuable player makes less than the average office worker in 
New York (an exaggeration, maybe, but Elena Delle Donne made around 120 thousand USD in her MVP year with little extra 
endorsement money or exposure). An NBA rookie signs one contract and has enough wealth to live comfortably off for 
the rest of his life, often for generations; a WNBA prospect is encouraged to continue her academic studies to 
ensure that she is able to get a good job after her career is over. Now, granted, the league revenue is much 
much less than that of the NBA, but many questions still remain that this paper hopes to address: Are the players 
paid fairly compared to league revenue? Are they paid fairly compared to NBA players based on basketball production? 
How can this change?
\par
There have been numerous studies surrounding that of the NBA and the WNBA and the differences in salaries.
For example, one such paper authored by Elle Baker discusses the intra-league salary distribution and
how WNBA players, despite getting paid far less, have much less "salary inequality"
\cite{baker2020comparison}. Another argues for an increase in salaries in the WNBA (a cause everyone
should be able to get behind) by citing the league's growth in fans and revenues \cite{ettienne2019s}.
There seems to be a lack of specific adjusted salary comparisons, however, and I hope to build off articles
like these to further the conversation.
\par
In this paper, I hope to compare not just salaries of WNBA vs NBA players, which is something that has been done 
numerous times, but to go further in depth and examine salaries vs team salary caps, league revenues, games played, 
and statistical performances for each player. The goal is to draw conclusions on if there are certain statistics in 
the NBA that result in higher adjusted salaries than the WNBA and vice versa, what the correlation is between them, 
and determine superstar salary proportions (amongst similar questions). These are related, yet more in depth questions 
to the ones posed by Elle Baker.
\par
The paper will begin by discussing the datasets being analyzed and how they were cleaned and prepped for analysis using 
Python scripts. In the next section, statistical analysis will be performed to try and answer some of the questions 
posed in this introduction. Following that will be a discussion on the results and hypotheses for best paths forward. 
Finally, there will be a conclusions section to wrap up the paper and review the results.

\paragraph{Data}
There are two primary data sets that I will be using to conduct the analysis in this paper. The first of these is a 
database of historical WNBA data dating from 2001-2020, collected by Neil Payne, sports editor at 538, and displayed 
on his GitHub account \cite{first}. Unlike most data sets, this one contains a multitude of advanced statistics for 
each player. The statistical categories are as follows:
\newline
\par
Player Name, Year, Age, Team, Team Games, Team Net Rating, Position, Games, Minutes Played, Minutes Played (as a percentage 
of team total minutes), Player Efficiency Rating, True Shooting \%, Free Throw Rating, Offensive Rebounding \%, Total 
Rebounding \%, Assist \%, Steal \%, Block \%, Turnover \%, Usage \%, Offensive Win Shares, Defensive Win Shares, 
Win Shares, Win Shares per 40 games, Composite Rating, and Wins Generated.
\newline
\par
Using another data set from Paine’s Github, a Python script called WNBA\_Cleanup.py was written and used to combine 
salary data from the two sources and filter out the desired candidate players from the data set. This paper will 
focus on comparing players from the 2018 NBA and WNBA seasons, so that filter was applied. The script also 
filtered out players who played less than 5 games and those that did not have salary data, so as to make the set as 
standardized as possible and remove outliers. The final data set after filtration and combination was called 
WNBA\_Combined.csv and has data (the aforementioned statistical categories) for 166 WNBA players from the 2018 season.
\par
A corresponding NBA dataset was found, this time belonging to Mustafa Baris Camli on the data-sharing site 
Kaggle \cite{nba}. This data set contains a list of all players from the NBA in the 2017-18 season, with similar 
advanced statistics to the WNBA dataset:
\newline
\par
Player Name, Position, Age, Team, Games Played, Minutes Played, Player Efficiency Rating, True Shooting \%, 
3 Point Attempt Rate, Free Throw Rate, Offensive Rebounding \%, Defensive Rebounding \%, Total Rebounding \%, 
Assist \%, Steal \%, Block \%, Turnover \%, Usage \%, Offensive Win Shares, Defensive Win Shares, Win Shares, 
Win Shares per 48 games, Offensive Box +/-, Defensive Box +/-, Box +/-, and VORP.
\newline
\par
The data set notably did not contain player salaries, so a second NBA data set was necessary. One was found on GitHub, 
created by Erik Gregory Webb, which contains the player salaries for every NBA player from 2000-2020 \cite{nba_salaries}. 
In a similar manner to the WNBA data, a Python script called NBA\_Cleanup.py was created to combine the data from the two 
datasets into one csv file. The data was then cleaned up by removing outliers (players playing under 5 games) and those 
that did not have salary data. The script was also used to combine data for players who played on multiple teams; 
the original data set would have players listed separately for each team they played on. The final data set after filtration 
and combination was called NBA\_Combined.csv and has data (the aforementioned statistical categories, plus player salaries) 
for 376 NBA players from the 2017-18 season.
\par
Finally, in order to compare the players accurately, the goal of checking their stats and salaries against league 
revenues must be achieved. As such, it was found that the NBA league revenue from 2018 was 8.01 billion USD \cite{NBA_Revenue} 
and the WNBA league revenue from the same year was 60 million USD \cite{WNBA_Revenue}.

\paragraph{Research Design and Methods}
Initially, we began by performing a series of paired sample t-tests to confirm some expectations from the data. We do 
not need to perform a test to determine if NBA players are paid more than WNBA players, as that is commonly known and 
very obvious. As (INSERT PAPER) discovered, in general WNBA players are actually paid more than NBA players proportional 
to league revenue. In order to confirm that the data for the 2018 seasons fit with this mold, an initial t-test was performed:
$$H_0 = \mu_0 \leq \mu_1, H_a = \mu_0 > \mu_1$$
where $\mu_0$ represents the average WNBA salary proportional to league revenue and $\mu_a$ represents the same for 
the NBA. The test was done using a significance level of  5\%, or an alpha value of 0.05, and degrees of freedom 
$n_1 + n_2 - 2 = 516$. By the t-statistic chart, it must be checked if the found t-statistic is greater than 1.646. 
Using a Python script, Comp\_t\_tests.py, it was found that our t-statistic was 16.56, far greater than 1.646, 
so it can safely be said with 95\% confidence that our data follows the trend found in (INSERT PAPER)
\par
The primary focus of the analysis section was to analyze correlation using linear regression between advanced statistics 
and salary. This was done once again using python scripts: LR\_Test\_WNBA and LR\_Test\_NBA, for the WNBA and NBA, 
respectively, were used to read the data from the cleaned datasets and conduct simple linear regression analysis on 
all the advanced stats against player salaries. Then, using the scipy and matplotlib Python packages, each comparison 
was graphed with a line of best fit and a correlation coefficient ($R^2$) was found. This correlation coefficient 
tells us how strong the correlation effect is between the statistic and salary, which will help us draw significantly 
more accurate conclusions from the data then if simply the slope was examined. There were twelve overlapping 
statistics between the WNBA and NBA datasets, and the correlation coefficients for these statistics in both leagues
 are graphed below:

\begin{tabular}{*5l}    \toprule
WNBA vs NBA Correlation & Coefficients For & Selected Advanced Statistics \\\midrule

Statistic 
&  CC: WNBA   & CC: NBA \\

 PER  &  0.019458  & 0.25444 \\
 FTr   &  0.001641   & 0.033096        \\       
 ORB\%   &  0.019717   & 0.000274       \\         
 TRB\%   &  0.000017   & 0.007801        \\
 AST\%   &  0.041707   & 0.110984         \\
 STL\%  &  0.003666   & 0.006126        \\    
 BLK\%  & 0.000444    & 0.000141      \\
 TOV\%  & 0.010470   & 0.004207     \\
 USG\%   & 0.033326    & 0.202527        \\
 OWS   & 0.117693    & 0.337278        \\
 DWS  & 0.151183    & 0.234252         \\
 WS   &  0.153999  & 0.364165                  \\
 (WS/40)/(WS/48)   &  0.047009   & 0.151640        \\
 \hline
\end{tabular}
\newline
\par
The graphs for the strongest correlation in both leagues (both the win shares statistic) are shown below:
\newline
\includegraphics[width=.85\textwidth]{NBA_WS_Graph.png}
\newline
\includegraphics[width=.85\textwidth]{WNBA_WS_Graph.png}

\par
Finally, within the Comp\_t\_tests.py Python script, analysis was conducted on the salaries of the top players in 
each league. Instead of looking at simple statistics, work that has been done in (DiCicio), this method used the 
advanced statistics from the data set. Six statistics were selected: Minutes Played (MP), Player Efficiency Rating 
(PER), Free Throw rate (FTr), Offensive Win Shares (OWS), Defensive Win Shares (DWS), and Win Shares (WS). 
The script was written to sort the datasets by these statistics, and the top 10 players in each league in these 
statistics were extracted. Then the salaries were compared to each respective league's median salary, and the 
average of these top 10 players relations to the average were compiled. The entries in the graph can be interpreted 
as follows: If the value is greater than 1, than the top 10 players in that statistic are paid the value as a factor 
above league average. If the value is less than 1, than the top 10 players in that statistic are paid the value as a 
factor below league average. The graph is below:
\vspace*{1cm}
\begin{tabular}{*5l} 
\toprule
How Much Compared to Avg & the Top 10 Players Are Paid & For Advanced Stats \\\midrule

Statistic &  WNBA  & NBA \\

 MP  &  1.445  & 5.353 \\
 PER   &  1.254   & 5.608        \\       
 FTr   &  1.087   & 2.053       \\         
 OWS   & 1.445    & 5.666        \\
 DWS  & 1.435    & 3.462       \\
 WS   &  1.530  & 5.446                 \\
 \hline
\end{tabular}

\paragraph{Discussion}
The data found from the initial t-tests supported what had been found in the current literature. It was confirmed that in 
our 2018 data WNBA players are paid significantly less than NBA players, a known fact. It was also confirmed with a very 
high degree of certainty that WNBA players are paid more than NBA players proportional to league revenue (around 60 
million for the WNBA, 8 billion for the NBA). This confirms the findings of (DiCicio), but was found to be particularly 
interesting and is in general counterintuitive to the narrative surrounding pay in women’s basketball. It also changes 
the conversation on how the pay gap can be narrowed.
\par
The main focus of the research was the regression analysis. Simple linear regression was used to create $R^2$ 
correlation coefficients for every advanced statistic in the WNBA and NBA datasets compared to player salaries. 
Since the base salary and proportional salary in the leagues are so different (as shown in the aforementioned t-tests), 
it seemed most appropriate to simply see which statistics correlated the strongest to salary increases. Table 1 shows 
the selection of correlation coefficients for the thirteen matching statistics between the two league data sets.
\par
Surprisingly, the correlation coefficients for almost every statistic was very low. For certain less consequential 
statistics (like Free Throw rate) this discovery was not shocking, but for more relatively important ones like Win 
Shares / Minutes it was interesting to see such low correlation between performance and pay. That being said, of the 
most important statistics - Player Efficiency Rating and Win Shares (including the Offensive and Defensive Win Shares 
sub statistics) - the NBA saw a significant increase in correlation coefficient, with all being above 0.23. The WNBA, 
on the flipside, did not see such an increase: Player Efficiency Rating had an extremely small coefficient, and while 
WS, OWS and DWS were all higher than other statistics, they still capped out at 0.15, lower than 6 of the NBA statistics 
coefficients. The graphs for the two highest correlations, Win Shares, are shown, and it can be visibly seen that higher
 performance in the NBA leads to higher pay more often than in the women’s counterpart league.
\par
Finally, some data manipulation was done to find a more simple statistic to try and determine how excellence in a statistic 
impacts pay. As Table 2 shows, there is a much clearer trend in performing in the top 10 in a statistic and a pay increase. 
As the WNBA pay spread is much smaller than the NBA’s it is expected that there will be less of an increase. An encouraging 
sign is found in this table: finishing top 10 in every statistic (with the exception of Free Throw rate, the least 
consequential of the statistics) led to a significant pay increase in both leagues. The NBA consistently saw an increase 
of over 5 times league median salary, while the WNBA hovered around 1.5 times league median.
\par
None of the data found is incredibly shocking, but it adds important information to the current research on the subject. 
For the WNBA to be treated seriously by fans and potential athletes, they must begin to move closer to the NBA formula 
of greater correlation of performance to player paychecks. The WNBA is obviously limited by their revenue, but the product 
is there; women’s basketball is just as enjoyable to watch as mens, but the players remain treated so extremely differently. 
There is money to be had in women’s basketball, as evidenced by several of the collegiate players at the University of 
Connecticut making over \$1 million per year in endorsement deals under the new NCAA Name Image Likeness rules. To see 
such insignificant correlation between a statistic like Player Efficiency Rating and pay in the WNBA vs the NBA is eye
 opening to the gender sport inequality.

\paragraph{Conclusion}
The motivation of this paper was to try and determine what goes into the pay gap in men’s and women’s basketball 
and primarily to see if the salary models were similar (just dramatically scaled down in the much poorer league). 
It was confirmed that the data for the 2018 season fit with precedent in terms of NBA salaries being much higher 
and WNBA salaries being higher proportional to league revenue. The regression analysis revealed that the 
correlation coefficients for NBA salaries to selected impactful advanced statistics was in general much higher 
than that of the WNBA, revealing that the aforementioned salary models were dramatically different. It was, 
however, determined that finishing in the top 10 in said statistics resulted in pay increases over league median, 
however the NBA’s pay increases were significantly more due to their much larger spread in salaries. 
\par
Next steps from this data involve researching what goes into league revenue and attempting to determine why it 
has never been profitable. It is important to acknowledge one last time that the WNBA is hamstrung by this lack 
of revenue, but the contract structures reinforce male dominant stereotypes in an already masculine world with a 
widespread gender pay gap in nearly every industry. Ultimately, the NBA players get rewarded much more for 
performance than the WNBA, which may motivate them to improve more and bring more excitement and popularity to the league. 
With less monetary motivation to perform, the WNBA continues to sell itself well short of its potential and does 
not give the players the respect that they deserve.

\bibliography{../Proposal/refs}
\bibliographystyle{chicago}

\end{document}